---
title: "A summary of the tweets generated at the 2019 CEFP workshop"
author: "Rayna M Harris"
output: md_document
---

```{r setup, echo=FALSE, warning=F, message=F}
## load packages
library(rtweet) # for fetching twitter data
library(dplyr) # for filter and select
library(tidyr) # for gather function
library(knitr) # for pretty tables
library(ggplot2) # for plotting
library(cowplot) # for multiplots

# set default code output to False
knitr::opts_chunk$set(echo = F)

# set figure directory
knitr::opts_chunk$set(fig.path = 'rtweets_CEFP2019_figs/')

```


```{r fetch tweets, message=F}
## first, follow steps in <https://rtweet.info/> to setup API and relevant tokens

## search for 18000 tweets using the rstats hashtag
rt <- search_tweets(
  "#CEFP2019", n = 18000, include_rts = FALSE
)
```

```{r fix tweet}
# fix stephanie butland tweet
rt[66, "text"] <- "I'm the luckiest community manager! Mentoring this week for 2019 cohort of AAAS Community Engagement Fellows. I learn from @rOpenSci community, @_inundata @sckottie @ma_salmon @opencpu @RLadiesGlobal @bffo &amp; #CEFP2017 peers. Taking all the goodness back to them #CEFP2019. https://t.co/ybvfbV8RTX"
```


### Basic summary

There were a total of **`r nrow(rt)`** tweets with the #CEFP2019 hashtag. These tweets were authored by **`r length(unique(rt$screen_name))`** different particpants. 

```{r topusers}
top_users <- rt %>%
    group_by(screen_name) %>%
    summarize(total_tweets = n()) 

ggplot(data = top_users) + 
  geom_bar(aes(x = reorder(screen_name, total_tweets), y = total_tweets), 
           stat = "identity", fill = "#505050") +
  geom_text(aes(label = total_tweets, y = total_tweets, x = screen_name),  
            hjust=0, size=3) +
  coord_flip() +
  scale_fill_manual(values = wes_palette("Zissou1")[c(1, 3)]) +
  theme_minimal() +
  theme(axis.text = element_text(size = 8),
  legend.text = element_text(size = 8)) + 
  labs(x = "User", y = NULL,
    title = "Total number of #CEFP2019 tweets by user",
    caption = "\nSource: Data collected from Twitter's REST API via rtweet")
```

```{r tweet_timeline}
# tweet timeline
ts_plot(rt, "2 hours") +
  ggplot2::theme_minimal() +
  ggplot2::theme(plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Frequency of #CEFP2019 tweets",
    subtitle = "Total tweet counts aggregated using two-hour intervals",
    caption = "\nSource: Data collected from Twitter's REST API via rtweet"
  )

```


### The 5 most favorited tweets

```{r top5favs, results='asis'}
top5favs <- rt %>%
  top_n(5, favorite_count) %>%
  arrange(desc(favorite_count)) %>%
  select(screen_name, text)
kable(top5favs)
```

### The 5 most retweeted tweets

```{r topretweets, results='asis'}
topretweets <- rt %>%
  top_n(5, retweet_count) %>%
  arrange(desc(retweet_count)) %>%
  select(screen_name, text, retweet_count)
kable(topretweets)
```

### Which tweeters had the highest impact?

```{r}
impact <- rt %>%
    group_by(screen_name) %>%
    summarize(n_tweets = n(),
              n_fav = sum(favorite_count),
              n_rt =  sum(retweet_count),
              mean_fav = round(mean(favorite_count), digits = 2),
              mean_rt = round(mean(retweet_count), digits = 2)) %>%
    filter(n_tweets >=  3) %>%
    arrange(-n_tweets)
#kable(impact)
```

```{r mostfav}
most_fav <- impact %>%
    top_n(20, n_fav)
most_fav$screen_name <- reorder(most_fav$screen_name,
                              most_fav$mean_fav,
                              sort)
a <- ggplot(most_fav) +
    geom_bar(aes(x = reorder(screen_name, n_fav), y = n_fav),
             stat = "identity", fill = "#0C2939") +
    coord_flip() +  
    geom_text(aes(label = n_fav, y = n_fav, x = screen_name),  
            hjust=1, size = 2.5, color = "#E1E9E8") +
    theme_minimal() +
    theme(axis.text = element_text(size = 8),
          legend.text = element_text(size = 8)) +
  labs(x = "User", y = NULL,
    title = "Total #CEFP2019 favorites")

```

```{r meanfav}
mean_fav <- impact %>%
    top_n(20, mean_fav)
mean_fav$screen_name <- reorder(mean_fav$screen_name,
                              mean_fav$mean_fav,
                              sort)
b <- ggplot(mean_fav) +
    geom_bar(aes(x = reorder(screen_name, mean_fav), y = mean_fav),
             stat = "identity", fill = "#01509D") +
    coord_flip() +  
    geom_text(aes(label = mean_fav, y = mean_fav, x = screen_name),  
            hjust=1, size = 2.5, color = "#E1E9E8") +
    theme_minimal() +
    theme(axis.text = element_text(size = 8),
          legend.text = element_text(size = 8)) +
  labs(x = "User", y = NULL,
    title = "Mean #CEFP2019 favorites") 

plot_grid(a,b)

```

```{r totalretweet}
most_rt <- impact %>%
    top_n(20, n_rt)
most_rt$screen_name <- reorder(most_rt$screen_name,
                              most_rt$n_rt,
                              sort)
c <- ggplot(most_rt) +
    geom_bar(aes(x = reorder(screen_name, n_rt), y = n_rt),
             stat = "identity", fill = "#005067") +
    coord_flip() + 
    geom_text(aes(label = n_rt, y = n_rt, x = screen_name),  
            hjust=1, size = 2.5, color = "#E1E9E8") +
    theme_minimal() +
    theme(axis.text = element_text(size = 8),
          legend.text = element_text(size = 8)) +
  labs(x = "User", y = NULL,
    title = "Total #CEFP2019 retweets")
```


```{r meanretweet}
most_rt <- impact %>%
    top_n(20, mean_rt)
most_rt$screen_name <- reorder(most_rt$screen_name,
                              most_rt$mean_rt,
                              sort)
d <- ggplot(most_rt) +
    geom_bar(aes(x = reorder(screen_name, mean_rt), y = mean_rt),
             stat = "identity", fill = "#00819E") +
    coord_flip() + 
    geom_text(aes(label = mean_rt, y = mean_rt, x = screen_name),  
            hjust=1, size = 2, color = "#E1E9E8") +
    theme_minimal() +
    theme(axis.text = element_text(size = 8),
          legend.text = element_text(size = 8)) +
  labs(x = "User", y = NULL,
    title = "Mean #CEFP2019 retweets")

plot_grid(c,d)
```

### Acknowledgments

My [source code](https://github.com/raynamharris/cefp2019/blob/master/dataviz/rtweets_CEFP2019.Rmd) was adapted from François Michonneau's [code](https://github.com/fmichonneau/2018-carpentrycon-tweets/blob/master/index.Rmd) that he used to create this [blog post about twitter statistics from Carpentry Con 2018](https://carpentries.org/2018/06/carpentrycon-tweets). The [`rtweet` package](https://rtweet.info/) had excellent documentation. I used [Image Color Picker](https://imagecolorpicker.com/) to obtain hex color codes following the CEFP color palette.
