---
title: "A summary of the tweets generated at SBN 2019"
author: "Rayna M Harris"
output: html_document
---

```{r setup, echo=FALSE, warning=F, message=F}
## load packages
#install.packages("plotrix")
library(rtweet) # for fetching twitter data
library(dplyr) # for filter and select
library(tidyr) # for gather function
library(knitr) # for pretty tables
library(ggplot2) # for plotting
library("plotrix") # for std error
library("cowplot") # for doubble figures
library("ggtextures") 


# set default code output to False
knitr::opts_chunk$set(echo = F)

# set figure directory
knitr::opts_chunk$set(fig.path = 'rtweets_SBN2019_figs/')

```


```{r fetch tweets, message=F}
## first, follow steps in <https://rtweet.info/> to setup API and relevant tokens

## search for 18000 tweets using the rstats hashtag
rt <- search_tweets(
  "sbn2019 OR sbn19 OR sbntweets ", n = 1000, include_rts = FALSE, type = "mixed"
)
```


### Basic summary

There were a total of **`r nrow(rt)`** tweets with the SBN hashtag. These tweets were authored by **`r length(unique(rt$screen_name))`** different particpants. 

```{r topusers}
top_users <- rt %>%
    group_by(screen_name) %>%
    summarize(total_tweets = n()) %>%
  filter(total_tweets > 5) %>%
  mutate(div6 = total_tweets / 6) %>%
  arrange(div6)


top_users$image <- c(
  "https://upload.wikimedia.org/wikipedia/commons/thumb/6/61/Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg/220px-Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg",
"https://upload.wikimedia.org/wikipedia/commons/thumb/6/61/Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg/220px-Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg",
"https://upload.wikimedia.org/wikipedia/commons/thumb/6/61/Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg/220px-Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg",
 "https://upload.wikimedia.org/wikipedia/commons/thumb/6/61/Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg/220px-Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg",
"https://upload.wikimedia.org/wikipedia/commons/thumb/6/61/Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg/220px-Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg",
"https://upload.wikimedia.org/wikipedia/commons/thumb/6/61/Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg/220px-Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg",
 "https://upload.wikimedia.org/wikipedia/commons/thumb/6/61/Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg/220px-Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg",
"https://upload.wikimedia.org/wikipedia/commons/thumb/6/61/Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg/220px-Taeniopygia_guttata_-_Bushell%27s_Lagoon.jpg",
"https://sites.google.com/site/ingridcaboucaud/_/rsrc/1417024312307/zebra-finch/DSCN0218.png?height=274&width=320",
 "https://sites.google.com/site/ingridcaboucaud/_/rsrc/1417024312307/zebra-finch/DSCN0218.png?height=274&width=320",
"https://sites.google.com/site/ingridcaboucaud/_/rsrc/1417024312307/zebra-finch/DSCN0218.png?height=274&width=320",
"https://sites.google.com/site/ingridcaboucaud/_/rsrc/1417024312307/zebra-finch/DSCN0218.png?height=274&width=320",
 "https://sites.google.com/site/ingridcaboucaud/_/rsrc/1417024312307/zebra-finch/DSCN0218.png?height=274&width=320",
"https://encrypted-tbn0.gstatic.com/images?q=tbn:ANd9GcQ4XwyZazHHPn2MrW2ZzrSs7p5hfJyHbwGdA-9o_SA0nPpb59d9",
 "https://encrypted-tbn0.gstatic.com/images?q=tbn:ANd9GcQ4XwyZazHHPn2MrW2ZzrSs7p5hfJyHbwGdA-9o_SA0nPpb59d9"
)


head(top_users)

ggplot(top_users, aes(reorder(screen_name, total_tweets), total_tweets, image = image)) +
  geom_textured_col(img_width = unit(0.1, "null")) +
  geom_text(aes(label = total_tweets, y = total_tweets, x = screen_name),  
            hjust= 0.1, size = 2.5) +
  coord_flip() +
  #scale_fill_manual(values = wes_palette("Zissou1")[c(1, 3)]) +
  theme_bw() +
  theme(axis.text = element_text(size = 8),
  legend.text = element_text(size = 8)) + 
  labs(x = "User", y = NULL,
    title = "Who live tweeting #SBN2019 OR #SBN19 OR @SBNtTweets?",
    subtitle = "Total number of tweets at SBN 2019",
    caption = "\nSource: Data collected from Twitter's REST API via rtweet using ")

ggplot(top_users, aes(reorder(screen_name, total_tweets), div6, image = image)) +
  geom_textured_col(img_width = unit(0.1, "null")) +
  geom_text(aes(label = total_tweets, y = total_tweets, x = screen_name),  
            hjust= 0.1, size = 2.5) +
  coord_flip() +
  #scale_fill_manual(values = wes_palette("Zissou1")[c(1, 3)]) +
  theme_bw() +
  theme(axis.text = element_text(size = 8),
  legend.text = element_text(size = 8)) + 
  labs(x = "User", y = NULL,
    title = "Who live tweeting #SBN2019 OR #SBN19 OR @SBNtTweets?",
    subtitle = "Total number of tweets at SBN 2019",
    caption = "\nSource: Data collected from Twitter's REST API via rtweet using ")

```

```{r tweet_timeline}
# tweet timeline
ts_plot(rt, "2 hours") +
  ggplot2::theme_minimal() +
  ggplot2::theme(plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Frequency of SBN tweets",
    subtitle = "Total tweet counts aggregated using two-hour intervals",
    caption = "\nSource: Data collected from Twitter's REST API via rtweet"
  )

```



```{r top5favs, results='asis'}
top5favs <- rt %>%
  top_n(5, favorite_count) %>%
  arrange(desc(favorite_count)) %>%
  select(screen_name, text, favorite_count)
head(top5favs)
```


```{r topretweets, results='asis'}
topretweets <- rt %>%
  top_n(5, retweet_count) %>%
  arrange(desc(retweet_count)) %>%
  select(screen_name, text, retweet_count)
head(topretweets)
```

### Which tweeters had the highest impact?

```{r}
impact <- rt %>%
    group_by(screen_name) %>%
    summarize(n_tweets = n(),
              n_fav = sum(favorite_count),
              n_rt =  sum(retweet_count),
              mean_fav = round(mean(favorite_count), digits = 2),
              mean_rt = round(mean(retweet_count), digits = 2),
              se_fav = std.error(favorite_count),
              se_rt = std.error(retweet_count)) %>%
    filter(n_tweets >=  5) %>%
    arrange(-n_tweets)
head(impact)
#head(impact)
```

```{r retweet}

mean_rt <- impact %>%
    top_n(20, mean_rt)
mean_rt$screen_name <- reorder(mean_rt$screen_name,
                              mean_rt$mean_rt,
                              sort)



b2 <- mean_rt  %>% 
  ggplot(aes(x= screen_name, y = mean_rt, 
             stat = "identity")) + 
    geom_errorbar(aes(ymin=mean_rt-se_rt, ymax=mean_rt+se_rt), width=.1) +
    geom_point(size = 2) +
    coord_flip() + 
    geom_text(aes(label = mean_rt, y = mean_rt, x = screen_name),  
            hjust=1, size = 2.5, color = "#E1E9E8") +
    theme_minimal() +
    theme(axis.text = element_text(size = 8),
          legend.text = element_text(size = 8)) +
  labs(x = NULL,
    subtitle = "Mean and total retweets at SBN",   
    y = "Mean retweets") +
    geom_text(aes(label = n_tweets), hjust = -2)

```

```{r favorites}
most_fav <- impact %>%
    top_n(20, n_fav)
most_fav$screen_name <- reorder(most_fav$screen_name,
                              most_fav$mean_fav,
                              sort)



mean_rt <- impact %>%
    top_n(20, mean_rt)
mean_rt$screen_name <- reorder(mean_rt$screen_name,
                              mean_rt$mean_fav,
                              sort)



d2 <- mean_rt  %>% 
  ggplot(aes( x = screen_name, y = mean_fav, 
             stat = "identity")) + 
    geom_errorbar(aes(ymin=mean_fav-se_fav, ymax=mean_fav+se_fav), width=.1) +
    geom_point(size = 2) +
    coord_flip() + 
    theme_minimal() +
    theme(axis.text = element_text(size = 8),
          legend.text = element_text(size = 8)) +
  labs(x = NULL, y = "Mean favorites",
    subtitle = "Mean and total favorites at SBN") +
  geom_text(aes(label = n_tweets), hjust = -2)
 
#d2

plot_grid(b2, d2)
```



### Acknowledgments

My [source code](https://github.com/raynamharris/cefp2019/blob/master/dataviz/rtweets_SBN2019.Rmd) was adapted from François Michonneau's [code](https://github.com/fmichonneau/2018-carpentrycon-tweets/blob/master/index.Rmd) that he used to create this [blog post about twitter statistics from Carpentry Con 2018](https://carpentries.org/2018/06/carpentrycon-tweets). The [`rtweet` package](https://rtweet.info/) had excellent documentation.